### 2021-02-23 architecture.tex
Shave off content by commenting out to reduce page count to 15

The recommender consists of modules that perform specific subtasks,
each of which is described in detail in one of the following sections (see also \cref{fig:architecture}):
\begin{itemize}
\item \Cref{sec:graphifier}: Graphifier converts the input \gls{cnf} problem into a problem graph.
\item \Cref{sec:gcn}: \Gls{gcn} converts a problem graph into symbol embeddings.
\todo{MS: Kdyz embeddingy nebudou v preliminaries, je asi OK ctenare, ktery by o nich neslysel, zde trochu vydesit.
Ale v te prislusne podsekci tim zas zacit.}
\item \Cref{sec:output}: An output function \cite{Zhou2018} converts the symbol embeddings into a vector of symbol costs.
\todo{MS: ``output function \cite{Zhou2018}'' my nic nerika, mel bych to znat?\\
FB: Zhou2018 je review GNNs. Output function je termin, ktery se tam pouziva. Ta citace je nejspis tady nepotrebna.}
\item \Cref{sec:ranking}: Sorting converts the symbol costs into a symbol precedence.
It is only used in generating mode.
\item \Cref{sec:training}: Loss function converts the symbol costs and a pair of precedences into a loss value.
It is only used in training mode.
\end{itemize}
\todo{MS: tohle zatim chapu jako predbezny nastrel. Leda by uz pred tim bylo vysvetleno:
embedding, \gls{gcn}/convolution layers, ... \\
FB: Staci predestrit, ze detaily vysvetlim pozdeji? MS: asi jo :)}

\begin{figure}[h]
\caption{Recommender architecture overview.
When recommending a precedence, the input is problem $P$ and the output is precedence $\pi^*$.
When training on an example, the input is problem $P$ and precedences $\pi$ and $\rho$,
and the output is the loss value.}
\label{fig:architecture}
\centering
\digraph[scale=0.4]{ArchitectureOverview}{
graph [splines=ortho,ranksep=0.2];
node [shape=box, fontsize=20];
edge [fontsize=20];

Problem [label="CNF problem P"];
Graphifier [style=rounded, label="Graphifier"];
g [label="Graph"];
GCN [style=rounded, label="GCN"];
SymbolEmbedding [label="Symbol embeddings"];
SymbolCostModel [style=rounded, label="Projection"];
SymbolCost [label="Symbol costs"];
Sort [style=rounded, label="Sort"];
Precedence [label=<Output precedence &pi;*>];

Problem -> Graphifier -> g -> GCN -> SymbolEmbedding -> SymbolCostModel -> SymbolCost -> Sort -> Precedence;

pi [label=<Precedence &pi;>];
rho [label=<Precedence &rho;>];
LossFunction [style=rounded, label="Loss function"];
Loss [label="Loss value"];

pi -> LossFunction;
rho -> LossFunction;
SymbolCost -> LossFunction;
LossFunction -> Loss;
}

\iffalse

\usetikzlibrary{shapes}
\tikzstyle{object} = [rectangle, draw]
\tikzstyle{input} = [ellipse, draw]
\tikzstyle{output} = [ellipse, draw]

\begin{tikzpicture}[node distance = 0.5 and 1, ->]
% https://tex.stackexchange.com/a/332796/202639

\node (problem) [input] {\gls{cnf} problem $P$};
\node (symbol embeddings) [object] [below=of problem] {Symbol embeddings};
\node (symbol costs) [object] [below=of symbol embeddings] {Symbol costs};
\node (symbol precedence) [output] [below=of symbol costs] {Symbol precedence};

\draw (problem) to node [left] {GCN} (symbol embeddings);
\draw (symbol embeddings) to node [left] {MLP} (symbol costs);
\draw (symbol costs) to node [left] {Order symbols by their costs} (symbol precedence);

\node (PrecBetter) [input] [right=of problem] {Precedence $\PrecBetter$};
\node (PrecBetterInv) [object] [below=of PrecBetter] {$\inv{\PrecBetter}$};
\draw (PrecBetter) to node [right] {Invert} (PrecBetterInv);

\node (PrecWorse) [input] [right=of PrecBetter] {Precedence $\PrecWorse$};
\node (PrecWorseInv) [object] [below=of PrecWorse] {$\inv{\PrecWorse}$};
\draw (PrecWorse) to node [right] {Invert} (PrecWorseInv);

\node (PrecDiff) [object] [right=2 of symbol costs] {$\inv{\PrecWorse} - \inv{\PrecBetter}$};
\node (PrecDiffNormalized) [object] [below=of PrecDiff] {Normalized};
\node (PrecPairCost) [object] [below=of PrecDiffNormalized] {Precedence pair cost};
\node (loss) [output] [below=of PrecPairCost] {Loss};

\draw (PrecBetterInv) to (PrecDiff);
\draw (PrecWorseInv) to (PrecDiff);
\draw (PrecDiff) to node [right] {Normalize} (PrecDiffNormalized);
\draw (PrecDiffNormalized) to (PrecPairCost);
\draw (PrecPairCost) to (loss);

\draw (symbol costs) to (PrecPairCost);

\end{tikzpicture}

\fi

\end{figure}
\todo{FB: Improve the diagram. Consider using tikz. See the disabled tikz figure in the source file.}

@@

As the first step of the recommender processing pipeline,
the input problem is converted from a \gls{cnf} representation
to a \emph{heterogeneous (directed) graph} \cite{Zhou2018}.
% Zhou uses the term "heterogeneous graph".
%\footnote{Data mining literature often uses the term \gls{hin} \cite{Shi2015} for heterogeneous graphs.
%We prefer to conform to the terminology common in literature studying \glspl{gnn}.}
Each of the nodes of the graph is labeled with a node type,
and each edge is labeled with an edge type,
defining the heterogeneous nature of the graph.
Each node corresponds to one of the elements that constitute the \gls{cnf} formula,
such as a clause, an atom or a predicate symbol.
Each such category of elements corresponds to one node type.
The edges represent the (oriented) relations between the elements,
for example the incidence relation between a clause and one of its (literals') atoms,
or the relation between an atom and its predicate symbol.
$\mathcal{R}$ denotes the set of all relations in the graph.
\Cref{fig:CnfSchema} shows the types of nodes and edges used in our graph representation.
\Cref{fig:GcnExample} shows an example of a graph representation of a simple problem.

@@

\begin{figure}[h]
\caption{CNF network schema}
\label{fig:CnfSchema}
\centering
\tikzstyle{token} = [rectangle, draw]

\begin{tikzpicture}[node distance = 1 and 2, ->]
% https://tex.stackexchange.com/a/332796/202639

% Node and edge types:
% https://docs.google.com/spreadsheets/d/1PCPHEgk6vLxpdpcvB_PGoLx7p4DID6WtvVWy2mDuv4A/edit?usp=sharing

\node (formula) [token] {\ntype{problem}};
% TODO: Consider removing the Formula nodes.
\node (clause) [token, below=of formula] {\ntype{clause}};
\node (atom) [token, below left=of clause] {\ntype{atom}};
\node (equality) [token, below right=of clause] {\ntype{equality}};
\node (predicate) [token, left=of atom] {\ntype{predicate}};
\node (argument) [token, below=of atom] {\ntype{argument}};
\node (term) [token, below=of argument] {\ntype{term}};
\node (function) [token, left=of term] {\ntype{function}};
\node (variable) [token, right=of term] {\ntype{variable}};

\draw (formula) to node [right] {\etype{contains}} (clause);
\draw (clause) to [bend right] node [above] {\epos{}} (atom);
\draw (clause) to [bend left] node [below] {\eneg{}} (atom);
\draw (clause) to [bend left] node [above] {\epos{}} (equality);
\draw (clause) to [bend right] node [below] {\eneg{}} (equality);
\draw (clause) to node [above] {\etype{binds}} (variable);
\draw (atom) to node [above] {\etype{atom\_applies}} (predicate);
\draw (atom) to node [left] {\etype{atom\_has}} (argument);
\draw (equality) to node {\etype{equalizes}} (term);
\draw (equality) to node {\etype{equalizes}} (variable);
\draw (argument) to [bend right] node [left] {\etype{is}} (term);
\draw (argument) to [loop left] node [left] {\etype{precedes}} (argument);
\draw (argument) to node [below] {\etype{is}} (variable);
\draw (term) to node [above] {\etype{term\_applies}} (function);
\draw (term) to [bend right] node [right] {\etype{term\_has}} (argument);

\end{tikzpicture}
\end{figure}
\todo{FB: Clean up the diagram.}

>>>>

The recommender consists of modules that perform specific subtasks,
each of which is described in detail in one of the following sections.
%\begin{itemize}
%\item \Cref{sec:graphifier}: Graphifier converts the input \gls{cnf} problem into a problem graph.
%\item \Cref{sec:gcn}: \Gls{gcn} converts a problem graph into symbol embeddings.
%\todo{MS: Kdyz embeddingy nebudou v preliminaries, je asi OK ctenare, ktery by o nich neslysel, zde trochu vydesit.
%Ale v te prislusne podsekci tim zas zacit.}
%\item \Cref{sec:output}: An output function \cite{Zhou2018} converts the symbol embeddings into a vector of symbol costs.
%\todo{MS: ``output function \cite{Zhou2018}'' my nic nerika, mel bych to znat?\\
%FB: Zhou2018 je review GNNs. Output function je termin, ktery se tam pouziva. Ta citace je nejspis tady nepotrebna.}
%\item \Cref{sec:ranking}: Sorting converts the symbol costs into a symbol precedence.
%It is only used in generating mode.
%\item \Cref{sec:training}: Loss function converts the symbol costs and a pair of precedences into a loss value.
%It is only used in training mode.
%\end{itemize}
%\todo{MS: tohle zatim chapu jako predbezny nastrel. Leda by uz pred tim bylo vysvetleno:
%embedding, \gls{gcn}/convolution layers, ... \\
%FB: Staci predestrit, ze detaily vysvetlim pozdeji? MS: asi jo :)}

%\begin{figure}[h]
%\caption{Recommender architecture overview.
%When recommending a precedence, the input is problem $P$ and the output is precedence $\pi^*$.
%When training on an example, the input is problem $P$ and precedences $\pi$ and $\rho$,
%and the output is the loss value.}
%\label{fig:architecture}
%\centering
%\digraph[scale=0.4]{ArchitectureOverview}{
%graph [splines=ortho,ranksep=0.2];
%node [shape=box, fontsize=20];
%edge [fontsize=20];
%
%Problem [label="CNF problem P"];
%Graphifier [style=rounded, label="Graphifier"];
%g [label="Graph"];
%GCN [style=rounded, label="GCN"];
%SymbolEmbedding [label="Symbol embeddings"];
%SymbolCostModel [style=rounded, label="Projection"];
%SymbolCost [label="Symbol costs"];
%Sort [style=rounded, label="Sort"];
%Precedence [label=<Output precedence &pi;*>];
%
%Problem -> Graphifier -> g -> GCN -> SymbolEmbedding -> SymbolCostModel -> SymbolCost -> Sort -> Precedence;
%
%pi [label=<Precedence &pi;>];
%rho [label=<Precedence &rho;>];
%LossFunction [style=rounded, label="Loss function"];
%Loss [label="Loss value"];
%
%pi -> LossFunction;
%rho -> LossFunction;
%SymbolCost -> LossFunction;
%LossFunction -> Loss;
%}
%
%\iffalse
%
%\usetikzlibrary{shapes}
%\tikzstyle{object} = [rectangle, draw]
%\tikzstyle{input} = [ellipse, draw]
%\tikzstyle{output} = [ellipse, draw]
%
%\begin{tikzpicture}[node distance = 0.5 and 1, ->]
%% https://tex.stackexchange.com/a/332796/202639
%
%\node (problem) [input] {\gls{cnf} problem $P$};
%\node (symbol embeddings) [object] [below=of problem] {Symbol embeddings};
%\node (symbol costs) [object] [below=of symbol embeddings] {Symbol costs};
%\node (symbol precedence) [output] [below=of symbol costs] {Symbol precedence};
%
%\draw (problem) to node [left] {GCN} (symbol embeddings);
%\draw (symbol embeddings) to node [left] {MLP} (symbol costs);
%\draw (symbol costs) to node [left] {Order symbols by their costs} (symbol precedence);
%
%\node (PrecBetter) [input] [right=of problem] {Precedence $\PrecBetter$};
%\node (PrecBetterInv) [object] [below=of PrecBetter] {$\inv{\PrecBetter}$};
%\draw (PrecBetter) to node [right] {Invert} (PrecBetterInv);
%
%\node (PrecWorse) [input] [right=of PrecBetter] {Precedence $\PrecWorse$};
%\node (PrecWorseInv) [object] [below=of PrecWorse] {$\inv{\PrecWorse}$};
%\draw (PrecWorse) to node [right] {Invert} (PrecWorseInv);
%
%\node (PrecDiff) [object] [right=2 of symbol costs] {$\inv{\PrecWorse} - \inv{\PrecBetter}$};
%\node (PrecDiffNormalized) [object] [below=of PrecDiff] {Normalized};
%\node (PrecPairCost) [object] [below=of PrecDiffNormalized] {Precedence pair cost};
%\node (loss) [output] [below=of PrecPairCost] {Loss};
%
%\draw (PrecBetterInv) to (PrecDiff);
%\draw (PrecWorseInv) to (PrecDiff);
%\draw (PrecDiff) to node [right] {Normalize} (PrecDiffNormalized);
%\draw (PrecDiffNormalized) to (PrecPairCost);
%\draw (PrecPairCost) to (loss);
%
%\draw (symbol costs) to (PrecPairCost);
%
%\end{tikzpicture}
%
%\fi
%
%\end{figure}
%\todo{FB: Improve the diagram. Consider using tikz. See the disabled tikz figure in the source file.}

@@

As the first step of the recommender processing pipeline,
the input problem is converted from a \gls{cnf} representation
to a \emph{heterogeneous (directed) graph} \cite{Zhou2018}.
% Zhou uses the term "heterogeneous graph".
%\footnote{Data mining literature often uses the term \gls{hin} \cite{Shi2015} for heterogeneous graphs.
%We prefer to conform to the terminology common in literature studying \glspl{gnn}.}
Each of the nodes of the graph is labeled with a node type,
and each edge is labeled with an edge type,
defining the heterogeneous nature of the graph.
Each node corresponds to one of the elements that constitute the \gls{cnf} formula,
such as a clause, an atom or a predicate symbol.
Each such category of elements corresponds to one node type.
The edges represent the (oriented) relations between the elements,
for example the incidence relation between a clause and one of its (literals') atoms,
or the relation between an atom and its predicate symbol.
$\mathcal{R}$ denotes the set of all relations in the graph.
%\Cref{fig:CnfSchema} shows the types of nodes and edges used in our graph representation.
\Cref{fig:GcnExample} shows an example of a graph representation of a simple problem.

@@

%\begin{figure}[h]
%\caption{CNF network schema}
%\label{fig:CnfSchema}
%\centering
%\tikzstyle{token} = [rectangle, draw]
%
%\begin{tikzpicture}[node distance = 1 and 2, ->]
%% https://tex.stackexchange.com/a/332796/202639
%
%% Node and edge types:
%% https://docs.google.com/spreadsheets/d/1PCPHEgk6vLxpdpcvB_PGoLx7p4DID6WtvVWy2mDuv4A/edit?usp=sharing
%
%\node (formula) [token] {\ntype{problem}};
%% TODO: Consider removing the Formula nodes.
%\node (clause) [token, below=of formula] {\ntype{clause}};
%\node (atom) [token, below left=of clause] {\ntype{atom}};
%\node (equality) [token, below right=of clause] {\ntype{equality}};
%\node (predicate) [token, left=of atom] {\ntype{predicate}};
%\node (argument) [token, below=of atom] {\ntype{argument}};
%\node (term) [token, below=of argument] {\ntype{term}};
%\node (function) [token, left=of term] {\ntype{function}};
%\node (variable) [token, right=of term] {\ntype{variable}};
%
%\draw (formula) to node [right] {\etype{contains}} (clause);
%\draw (clause) to [bend right] node [above] {\epos{}} (atom);
%\draw (clause) to [bend left] node [below] {\eneg{}} (atom);
%\draw (clause) to [bend left] node [above] {\epos{}} (equality);
%\draw (clause) to [bend right] node [below] {\eneg{}} (equality);
%\draw (clause) to node [above] {\etype{binds}} (variable);
%\draw (atom) to node [above] {\etype{atom\_applies}} (predicate);
%\draw (atom) to node [left] {\etype{atom\_has}} (argument);
%\draw (equality) to node {\etype{equalizes}} (term);
%\draw (equality) to node {\etype{equalizes}} (variable);
%\draw (argument) to [bend right] node [left] {\etype{is}} (term);
%\draw (argument) to [loop left] node [left] {\etype{precedes}} (argument);
%\draw (argument) to node [below] {\etype{is}} (variable);
%\draw (term) to node [above] {\etype{term\_applies}} (function);
%\draw (term) to [bend right] node [right] {\etype{term\_has}} (argument);
%
%\end{tikzpicture}
%\end{figure}
%\todo{FB: Clean up the diagram.}

### 2021-02-23 evaluation.tex
Add a todo
Shave off content by commenting out to reduce page count to 15

The scripts that were used to generate the training data and to train and evaluate the recommender
are available at the following URL:
\url{https://github.com/filipbartek/vampire-ml/tree/6907ee35}

@@

The empirical evaluation was performed using a modified version of the \gls{atp} \Vampire{} 4.3.0 \cite{10.1007/978-3-642-39799-8_1}.
The prover was used to generate the training data and to evaluate the trained precedence recommender.
To generate the training data,
\Vampire{} was modified to output \gls{cnf} representations of the problems
and annotated problem signatures in a machine-readable format.
%in \gls{json} format and annotated problem signatures in \gls{csv} format.\todo{MS: Mentioning JSON or CSV seems to be too low level. ``programmer documatation'' material.}
For the evaluation of the precedences generated by the recommender,
\Vampire{} was modified to allow the user to supply explicit predicate and function symbol precedences for the proof search
(normally, the user only picks a precedence generation heuristic).
\todo[inline]{FB: Add a link to the modified \Vampire{}.}

@@

\Vampire{} was run with a time limit of 10 seconds.
To increase the potential impact of predicate precedences,
we used a simple \gls{tkbo} \cite{Ludwig2007,Kovacs2011}
that compares atoms according to the predicate precedence first,
Using the ordinary \gls{kbo} as a tie-breaker.
Ordinary \gls{kbo} was used to compare terms.
\todo[inline,author=FB]{Describe the detailed configuration in an appendix and reference the appendix here.}

@@

The input problems are assumed to be specified in the \gls{cnf} or the \gls{fof} fragment
of the \acrshort{tptp} language \cite{Sutcliffe2017}.
% http://www.tptp.org/TPTP/SyntaxBNF.html
\Gls{fof} problems are first converted into equisatisfiable \gls{cnf} problems by \Vampire{}.
New symbols may be introduced during preprocessing,
namely due to the Tseitin transformation and Skolemization.
We denote the preprocessed \gls{cnf} problem as $P = (\Sigma, \mathit{Cl})$,
where $\Sigma$ denotes the list of all predicate and function symbols of the problem,
and $\mathit{Cl}$ denotes a set of clauses over $\Sigma$ that constitute the problem.

@@

After the training finished,
we performed a final evaluation of the most promising intermediate trained model on the whole $\ProblemsVal$.
The model that manifested the best estimated solver performance was chosen as the most promising.
\todo{Watch out: This way we may overfit on the 1000 problems used for sample solver performance evaluation.}
\todo{MS: taky me to ted napadlo! Uplne oficialni metodika byva split na train,valid,test. 
Kdy se trenuje na train, hyperparametry (vcetne early stopping) se ladi vuci valid 
a pak se vyhodnocuje na test. Nekdy se pak valid rika devel a test se rika holdout.}

@@

To generate a precedence for a problem,
\todo{MS: tenhle odstavec mozna prehodit s predchozim?}
the recommender first converts the problem to a machine-friendly \gls{cnf} format,
then converts the \gls{cnf} to a graph,
then predicts symbol costs using the \gls{gcn} model
and finally orders the symbols by their costs to produce the precedence.
To simplify the experiment, the time limit is only imposed on the \Vampire{} run
and excludes the time taken by the recommender to generate the precedence.
When run with 2 threads,
the preprocessing of a single problem
takes at most at most \num[round-mode=places,round-precision=2]{1.262744486} seconds
for \SI{80}{\percent} of the problems
according to data measured on a random sample of \num{1000} problems.
% Analysis: https://docs.google.com/spreadsheets/d/1GujYNEtETpC3jk4iyENLjptv8mDmI5f1ZEhbmwtqnPM/edit?usp=sharing
% VML-715
\todo{MS: 1) single-core bych chapal jako default. tj netreba zminovat?
2) kterych 1000? To opet drazdi k otazkam. tj to bych asi taky vynechal, ale...
3) to maximum prece zalezi na tom, kterych 1000 to bylo.}
\todo{MS: Idea: kdyz je to linearni: co cas potrebny na recommending
vztahnout k casu klauzifikace? Jako rict, ze to trva x-krat dyl nez jen klauzifikovat?}

@@

\begin{table*}[h]
\caption{
Results of the evaluation of various predicate precedence heuristics on $\ProblemsVal$
($\card{\ProblemsVal} = 7648$).
Means and standard deviations over 5 runs are reported.
}
\label{tab:results}
\centering
\begin{tabular}{l|ll|rl}

Model & \multicolumn{2}{l}{Successes on $\ProblemsVal$} & \multicolumn{2}{l}{Improvement over baseline} \\
& Mean & Std & Absolute & Relative \\

@@

\texttt{vampire -lcm standard} &
\num{3823.0} &
\num[round-mode=places,round-precision=2]{3.40587727318528} &
%\SI[round-mode=places,round-precision=2]{49.9869247}{\percent} &
+53.4 &
\num[round-mode=places,round-precision=3]{1.014165959} \\
% VML-741

\texttt{vampire -lcm predicate} &
% Success rate: mean: 0.492887029, std: 0.000401412
% https://ui.neptune.ai/filipbartek/vampire-ml/e/VML-490
% /home/filip/projects/vampire-ml/vampire-ml/outputs/2021-02-09/12-13-43
% Row: 'val&graphified&solver_eval'
% Problems total: 7648
% Success rate: 0.492887029
\num{3769.6} &
\num{3.07} &
%\SI{49.29}{\percent} &
0.0 &
\num[round-mode=places,round-precision=3]{1.0} \\

\end{tabular}
\end{table*}
\todo{FB: Add links to Neptune runs?}
\todo{FB: Finalize results for function precedences.}
\todo{FB: Try to evaluate a recommender that uses both predicate and function models.}

@@

The loss is merely a proxy measure of the accuracy,
and the accuracy is merely a proxy measure of the solver performance.
During the training, the validation loss typically started increasing after approximately 50 epochs,
while both the train loss and validation accuracy continued improving.
A natural possible explanation is that ...
\todo{Finish.}
While a higher granularity in the solver evaluation would be necessary to draw a clear conclusion,
the trend in the solver performance seems mostly aligned with the validation accuracy.

\todo[inline,author=FB]{Plot some learning curves?}

>>>>

The scripts that were used to generate the training data and to train and evaluate the recommender
are available at the following URL:
\url{https://github.com/filipbartek/vampire-ml/tree/6907ee35}
\todo{FB: Tag the version and link to the tag.}
\todo{FB: Add some wrapper scripts and readme for ease of use.}

@@

The empirical evaluation was performed using a modified version of the \gls{atp} \Vampire{} 4.3.0 \cite{10.1007/978-3-642-39799-8_1}.
The prover was used to generate the training data and to evaluate the trained precedence recommender.
To generate the training data,
\Vampire{} was modified to output \gls{cnf} representations of the problems
and annotated problem signatures in a machine-readable format.
%in \gls{json} format and annotated problem signatures in \gls{csv} format.\todo{MS: Mentioning JSON or CSV seems to be too low level. ``programmer documatation'' material.}
For the evaluation of the precedences generated by the recommender,
\Vampire{} was modified to allow the user to supply explicit predicate and function symbol precedences for the proof search
(normally, the user only picks a precedence generation heuristic).
The modified version of \Vampire{} is available at the following URL:
\url{https://github.com/filipbartek/vampire/tree/3bbfa9e6}

@@

\Vampire{} was run with a time limit of 10 seconds.
To increase the potential impact of predicate precedences,
we used a simple \gls{tkbo} \cite{Ludwig2007,Kovacs2011}
that compares atoms according to the predicate precedence first,
using the regular \gls{kbo} to break ties between atoms
and to compare terms.
\todo[inline,author=FB]{Describe the detailed configuration in an appendix and reference the appendix here.}

@@

The input problems are assumed to be specified in the \gls{cnf} or the \gls{fof} fragment
of the \acrshort{tptp} language \cite{Sutcliffe2017}.
% http://www.tptp.org/TPTP/SyntaxBNF.html
\todo{FB: Mention TPTP version. Consider joining with section Problem set below.}
\Gls{fof} problems are first converted into equisatisfiable \gls{cnf} problems by \Vampire{}.
New symbols may be introduced during preprocessing,
namely due to the Tseitin transformation and Skolemization.
We denote the preprocessed \gls{cnf} problem as $P = (\Sigma, \mathit{Cl})$,
where $\Sigma$ denotes the list of all predicate and function symbols of the problem,
and $\mathit{Cl}$ denotes a set of clauses over $\Sigma$ that constitute the problem.

@@

After the training finished,
we performed a final evaluation of the most promising intermediate trained model on the whole $\ProblemsVal$.
The model that manifested the best estimated solver performance on a sample of \num{1000} validation problems
was chosen as the most promising.
\todo{FB: Watch out: This way we may overfit on the 1000 problems used for sample solver performance evaluation. We should leave out these 1000 problems from the final evaluation.}
\todo{MS: taky me to ted napadlo! Uplne oficialni metodika byva split na train,valid,test. 
Kdy se trenuje na train, hyperparametry (vcetne early stopping) se ladi vuci valid 
a pak se vyhodnocuje na test. Nekdy se pak valid rika devel a test se rika holdout.}

@@

To generate a precedence for a problem,
\todo{MS: tenhle odstavec mozna prehodit s predchozim?}
the recommender first converts the problem to a machine-friendly \gls{cnf} format,
then converts the \gls{cnf} to a graph,
then predicts symbol costs using the \gls{gcn} model
and finally orders the symbols by their costs to produce the precedence.
To simplify the experiment, the time limit of 10 seconds is only imposed on the \Vampire{} run
and excludes the time taken by the recommender to generate the precedence.
When run with 2 threads,
the preprocessing of a single problem
takes at most at most \num[round-mode=places,round-precision=2]{1.262744486} seconds
for \SI{80}{\percent} of the problems
by extrapolation from a sample of \num{1000} problems.
% Analysis: https://docs.google.com/spreadsheets/d/1GujYNEtETpC3jk4iyENLjptv8mDmI5f1ZEhbmwtqnPM/edit?usp=sharing
% VML-715
\todo{MS: 1) single-core bych chapal jako default. tj netreba zminovat?
2) kterych 1000? To opet drazdi k otazkam. tj to bych asi taky vynechal, ale...
3) to maximum prece zalezi na tom, kterych 1000 to bylo.}
\todo{MS: Idea: kdyz je to linearni: co cas potrebny na recommending
vztahnout k casu klauzifikace? Jako rict, ze to trva x-krat dyl nez jen klauzifikovat?}

@@

\begin{table*}[h]
\caption{
Results of the evaluation of symbol precedence heuristics based on various symbol cost models
on $\ProblemsVal$ ($\card{\ProblemsVal} = 7648$).
Means and standard deviations over 5 runs are reported.
The \gls{gcn} models are trained according to the description in \cref{sec:architecture,sec:evaluation}.
The simple model is the final model from \cite{DBLP:conf/cade/Bartek020}.
The models that use machine learning for the predicate precedence only
use the \texttt{frequency} heuristic for the function precedence, and vice versa.
The frequency models use the standard \texttt{frequency} heuristic implemented in \Vampire{}.
}
\label{tab:results}
\centering
\begin{tabular}{l|ll|rl}

Symbol cost model & \multicolumn{2}{l}{Successes on $\ProblemsVal$} & \multicolumn{2}{l}{Improvement over baseline} \\
& Mean & Std & Absolute & Relative \\

@@

%Frequency (regular \acrshort{kbo}) &
%\texttt{vampire -lcm standard} &
%\num{3823.0} &
%\num[round-mode=places,round-precision=2]{3.40587727318528} &
%\SI[round-mode=places,round-precision=2]{49.9869247}{\percent} &
%+53.4 &
%\num[round-mode=places,round-precision=3]{1.014165959} \\
% VML-741

Frequency &
%\texttt{vampire -lcm predicate} &
% Success rate: mean: 0.492887029, std: 0.000401412
% https://ui.neptune.ai/filipbartek/vampire-ml/e/VML-490
% /home/filip/projects/vampire-ml/vampire-ml/outputs/2021-02-09/12-13-43
% Row: 'val&graphified&solver_eval'
% Problems total: 7648
% Success rate: 0.492887029
\num{3769.6} &
\num{3.07} &
%\SI{49.29}{\percent} &
0.0 &
\num[round-mode=places,round-precision=3]{1.0} \\

\end{tabular}
\end{table*}
\todo{FB: Add links to Neptune runs?}

@@

The results show that the \gls{gcn}-based model outperforms the \texttt{frequency} heuristic by a significant margin.
Since the predicate precedence recommender was trained against randomly distributed function precedences,
it is expected to perform well irrespective of the function precedence heuristic it is combined with.
Combining the trained recommenders for predicate and function precedences manifests better performance
than any of the two in combination with the standard \texttt{frequency} heuristic,
outperforming the \texttt{frequency} heuristic by approximately \SI{4.8}{\percent}.

We have verified that the simple linear predicate precedence heuristic
trained in \cite{DBLP:conf/cade/Bartek020} outperforms the \texttt{frequency} heuristic.
Moreover, we have confirmed the conjecture that using a \gls{gnn} may yield an even better performance \cite{DBLP:conf/cade/Bartek020}.

%In the process of training,
%we tracked three metrics on the training and validation sets:
%loss, accuracy, and solver performance on a sample of 1000 problems.
%A typical training ran for approximately 500 epochs
%with training loss steadily decreasing and
%validation loss decreasing for the initial 50 epochs and then increasing.
%Both training and validation accuracy were steadily increasing (training accuracy significantly faster than validation accuracy),
%with validation loss entering a plateau around epoch 100.
%The solver performance approximately followed the dynamics of the accuracy on the respective problem set.
% Predicate: https://ui.neptune.ai/filipbartek/vampire-ml/e/VML-540/charts
% Function: https://ui.neptune.ai/filipbartek/vampire-ml/e/VML-672/charts

\todo[inline]{FB: Discuss the phenomenon of increasing validation loss and accuracy.}
%A possible explanation of the discrepancy between the improving validation accuracy and deteriorating validation loss
%in the latter part of the training process
%is that the improvements of training loss generalize to improvements of loss on some, but not all of the validation examples.
%
%is that due to the limited generalization from training to validation set,
%some improvements of training loss come at the cost of an increase of loss on validation examples that are already misclassified
%(assigned a negative score).
%Such change increases the validation loss but does not affect the validation accuracy.
%At the same time there are validation examples whose loss is decreasing,

### 2021-02-23 main.tex
Shave off content by commenting out to reduce page count to 15

A more detailed analysis of the training process might yield insights into how the performance generalizes
from the training set to the validation set
and from proxy metrics, such as the loss and the accuracy, to the final solver performance.
Ultimately, this might reveal modifications that make the training more efficient and the final recommender more performant.

The training examples take the form of pairs of precedences,
where one of the precedences is preferable over the other.
In this work, the amount by which the quality of the precedences differs is ignored.
Taking the exact or estimated numbers of saturation loop iterations into account
might help the recommender to prioritize the more significant gains.

Applying techniques of reinforcement learning to focus the training data on the cases that yield high performance gains
could further improve the recommender performance.

We restricted our experiments to problems whose graph representation does not exceed \num{100000} nodes.
Evaluating the recommender on larger problems would be an interesting test of generalization.
Training on larger problems may be facilitated by employing a lossy graphification.

We performed all experiments with a fixed \gls{atp} \Vampire{} and a fixed configuration
we consider to be a reasonable baseline.
Introducing other prover configurations (namely other \glspl{sot})
or other provers could make the resulting recommender more robust
and possibly help extract more general knowledge about the structure of good symbol precedences.

Finally, the approach outlined in \cref{sec:architecture}
could be adjusted to solve tasks on \gls{fol} formulas other than recommending symbol precedences,
such as premise selection of given clause selection.
On the other hand, we could try to solve other tasks that involve generating permutations of arbitrary length.
\todo{FB: TSP? Probably not because it is too combinatorial.}

%KBO is a also parametrized by weights (along with the precedence). Could we learn those as well? (Jointly?)

>>>>

%A more detailed analysis of the training process might yield insights into how the performance generalizes
%from the training set to the validation set
%and from proxy metrics, such as the loss and the accuracy, to the final solver performance.
%Ultimately, this might reveal modifications that make the training more efficient and the final recommender more performant.
%
%The training examples take the form of pairs of precedences,
%where one of the precedences is preferable over the other.
%In this work, the amount by which the quality of the precedences differs is ignored.
%Taking the exact or estimated numbers of saturation loop iterations into account
%might help the recommender to prioritize the more significant gains.
%
%Applying techniques of reinforcement learning to focus the training data on the cases that yield high performance gains
%could further improve the recommender performance.
%
%We restricted our experiments to problems whose graph representation does not exceed \num{100000} nodes.
%Evaluating the recommender on larger problems would be an interesting test of generalization.
%Training on larger problems may be facilitated by employing a lossy graphification.
%
%We performed all experiments with a fixed \gls{atp} \Vampire{} and a fixed configuration
%we consider to be a reasonable baseline.
%Introducing other prover configurations (namely other \glspl{sot})
%or other provers could make the resulting recommender more robust
%and possibly help extract more general knowledge about the structure of good symbol precedences.
%
%Finally, the approach outlined in \cref{sec:architecture}
%could be adjusted to solve tasks on \gls{fol} formulas other than recommending symbol precedences,
%such as premise selection of given clause selection.
%On the other hand, we could try to solve other tasks that involve generating permutations of arbitrary length.
%\todo{FB: TSP? Probably not because it is too combinatorial.}
